\addcontentsline{toc}{chapter}{Acknowledgments}
\thispagestyle{empty}

\vspace*{20mm}

\begin{center}
    {\usekomafont{sectioning}\usekomafont{section} Acknowledgments}
\end{center}

\vspace{10mm}

I would like to express my deepest gratitude to my advisor, Kislaya Ravi, for the invaluable guidance, support, and encouragement throughout the course of this thesis project. 
Ravi's expertise, insights, and unwavering commitment to excellence have been instrumental in shaping this work. 
His mentorship has not only guided me through the complexities of this research but has also inspired me to think critically and approach challenges with determination. 
This thesis would not have been possible without Ravi's patience, constructive feedback, and constant encouragement, which motivated me to strive for improvement at every step.
Additionally, I would like to extend my early congratulations and best wishes to Ravi for his upcoming marriage next February. I wish him and his partner a lifetime of happiness, love, and success.

I am also profoundly grateful to my parents for the time and effort they have always dedicated to supporting me, both physically and spiritually. Their unwavering love, encouragement, and belief in me have been a constant source of strength and inspiration. I am truly blessed to have such loving and supportive parents who have always encouraged me to pursue my dreams and never give up, even during the hardest times.

Lastly, I would like to extend my heartfelt thanks to Dr. Tian Weikun, who has provided invaluable assistance during the writing of this thesis. Thank you for believing in me and for your unwavering support throughout this journey.

\cleardoublepage{}
